All Your Graphs Belong to Us

### 2012-07-18 doc/asplos13/implementation.tex
oops, fix build break
initial memory section

At the upper layer, Grappa implements asynchronous active messages
\cite{vonEicken92}. Each message consists of a function pointer, an
optional argument payload, and an optional data payload. When a task
sends a message, the message is copied to a send queue associated with
the message's destination and the task continues execution.

@@

Underneath the aggregation layer, Grappa uses the GASNet communication
library \cite{GASNet} to actually move data. All interprocess
communication, whether on or off a cluster node, is handled by the
GASNet library. GASNet is able to take advantage of many communication
mechanism, including Ethernet and Infiniband between nodes, as well as
shared memory within a node.

@@

remote access

heap

allocation

synchronization

consistency
>>>>>>> Stashed changes

>>>>

At the upper layer, Grappa implements asynchronous active messages
\TODO{cite vonEicken92}. Each message consists of a function pointer, an
optional argument payload, and an optional data payload. When a task
sends a message, the message is copied to a send queue associated with
the message's destination and the task continues execution.

@@

Underneath the aggregation layer, Grappa uses the GASNet communication
library \TODO{cite GASNet} to actually move data. All interprocess
communication, whether on or off a cluster node, is handled by the
GASNet library. GASNet is able to take advantage of many communication
mechanism, including Ethernet and Infiniband between nodes, as well as
shared memory within a node.

@@

Grappa implements a software distributed shared memory. This memory is
exposed through two address spaces; one optimized for global access to
node-local data, and another optimized for high aggregate random
access bandwidth to low locality data.

All global addresses in Grappa are 64-bit values. Each byte of global
memory is associated with a {\em home core}; this core is responsible
for allocating and modifying that memory. Since today's commodity
processors support a 48-bit virtual address space \TODO{cite?}, Grappa
uses the spare bits to distinguish which address space a global
address belongs to, and to represent the home core of a global
address.

The first global address space is the {\em two-dimensional} address
space. This uses a traditional PGAS addressing model, where each
address is a tuple of a rank in the job (or global process ID) and an
address in that process. The lower 48 bits of the address hold a
virtual address in the process. The top bit is set to indicate that
this a 2D address. This leaves 15 bits for network endpoint ID, which
limits our scalability to $2^15$ endpoints.

Any local data can be made accessible by other nodes in the system by
wrapping the address and node ID into a 2D global address. This
address can then be accessed with delegate or cache operations. At the
destination the address is converted into a canonical x86 address by
replacing the upper bits with the sign-exteded upper bit of the
virtual address.

The second global address space is the {\em linear} address space, so
named because nearby addresses can be treated as adjacent even if they
are stored on different nodes. Addresses in the linear address space
are partitioned across the cluster in a block-cyclic fashion. Choosing
the block size involves trading off sequential bandwidth agains
aggregate random access bandwidth. Smaller block sizes help spread
data across all the memory controllers in the cluster, but larger
block sizes allow the locality-optimized memory controllers to provide
increased sequential bandwidth. The block size is typically set to 64
bytes, or the size of a single hardware cache line, in order to
exploit spatial locality when available.

All linear addresses refer to data allocated from a global heap. The
heap metadata is stored on a single node. Currently all heap
operations serialize through this node; while this has been sufficient
for our benchmarks, parallel performance can be achieved through
combining \TODO{ cite MAMA and flat combining }. 2D addresses may
refer to memory allocated from a single processes' heap or from a
task's stack.

All global memory operations are implemented with active
messages. Requestors construct a descriptor holding the current
task/worker ID and space for a response, issue their request active
message to the node given by the request address, and suspend
themselves. When the request arrives at the address' home node, the
operation is performed and a response active message is sent. When the
response arrives at the requesting node, the descriptor is filled in
with the results of the operation and the requesting task is woken.

Delegate operations are used for syncronization. Their modifications
are always executed at the home core of their address. Each task may
have only one outstanding delegate operation. We limit delegate
operations to operate objects in the 2D address space or objects that
fit in a single block of the linear address space. No context switches
are allowed while the data it being modified. Given these
restrictions, we can ensure that delegate operations from multiple
requestors are always serialized through a single core in the system,
providing atomic semantics without using atomic operations.

Cache operations are used to exploit spatial or temporal
locality. They allow a single task to buffer a range of data, read or
modify it, and write the modified version back to its home node. All
acquires and releases are expressed explicitly by the
programmer. Cache operations may operate on data that spans multiple
blocks of the linear address space; the programmer provides the
starting address and length, and the cache issues as many active
messages as necessary to retrieve all necessary data.

All current cache operations are incoherent; two tasks may
concurrently modify the same block with their writes interfereing
unless protected by synchronization through delegate operations. While
this has been sufficient for our benchmarks, it is possible to embed
synchronization in the cache active messages and reduce the number of
network round trips necessary.
